Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural music skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-12 MasterThesis.tex
Updates from Overleaf

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{fadings}
\usetikzlibrary{matrix, positioning}
\usetikzlibrary{shapes,arrows,positioning}

@@

%Title parameters
\title{%
  Uncovering underlying musical content in the temporal domain \\
  \small Leveraging deep learning, inductive bias, and aural music skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks}

@@

%Introduction of the project
\input{introduction/introduction}
%Task definition
%%\input{task definition}
%Methods
\input{methods/methods}
%Results
\input{results/results}
%Discussion
\input{discussion/discussion}
%Future work
\input{future work/future work}
\newpage

>>>>

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,fadings,matrix, positioning,shapes,arrows,angles,quotes}

@@

%Title parameters
\title{%
  Uncovering underlying high-level musical content in the temporal domain \\
  \small Leveraging self-supervised deep neural networks, inductive bias, and aural music skills to learn music embeddings with applications to MIR downstream tasks}

@@

%Introduction of the project
\input{introduction/introduction}
%Task definition
%%\input{task definition}
%Methods
\input{methods/methods}
%Evaluation
\input{evaluation/evaluation}
%Results
\input{results/results}
%Discussion
\input{discussion/discussion}
%Future work
\input{future work/future work}
\newpage

### 2023-05-12 evaluation/subsections/beyond dsp.tex
Updates from Overleaf

\section{Beyond Digital Signal Processing (DSP)}

We argue that MIR research needs to incorporate a more balanced approach that considers the interdisciplinary nature of music and the importance of other domains beyond DSP.

\subsection{Sample MIR task}

@@

>>>>

\section{Music Structure Analysis as downstream task}

@@

\input{figures/neural networks/r128 sphere}

\section{SALAMI dataset}
\subsection{Overview}

SALAMI (Structural Analysis of Large Amounts of Music Information) \cite{Smith2011DESIGNANNOTATIONS} aims to conduct massive structural analyses of different types of music. The project covers diverse music, from Western pop to Indian classical and from live to studio recordings.

\subsubsection{Formal Analysis Definition}
Formal analysis, in this context, doesn't refer to classifying the piece into a formal type (e.g., sonata, song, or canon) or reducing it into its Ursatz as in Schenkerian analysis. Instead, it means organizing and dividing the piece into specific sections and understanding how they relate.

%%%%%%%%%%

\subsubsection{Approaches to Music Analysis}

The method proposed combines aspects of the following three approaches, separating the organization of instrumentation, musical material, and formal function. This allows for the analysis of a wide variety of music pieces in a consistent way using a constrained vocabulary. It's important to note that music structure is often hierarchical, and the appropriate timescale for analysis can be hard to determine. The proposed method includes a few markers to partly address this issue.

Perceptual Approach: This involves listening for prominent harmonic or rhythmic boundaries in the piece to segment and apply labels to sections based on their similarity. However, this approach might not reflect the different structural roles of similar-sounding sections.

Functional Approach: This involves dividing the song into its functional parts, like verses, choruses, intros, outros, etc. This might result in the same music being labeled differently based on its function or different musical ideas having the same functional label.

Transcription Approach: This involves recording musical parameters like chord patterns and instrumentation directly. While this method may be less subjective, it may not provide a comprehensive view of the piece's structure.

%%%%%%%

\subsubsection{SALAMI Annotation Labels}

The structural analysis of each piece involves partitioning it into sections and labeling these sections. There are three independent layers of labels: musical similarity, function, and instrumentation.

Musical Similarity: Labels (A, B, C, D, E, ...) represent large-scale musical phrases, ideas, or subjects differentiated based on rhythmic, melodic, or harmonic material. Each unique musical idea gets its label. Although the guide recommends limiting to five labels, more can be used. Every moment in the piece must be labeled.

'Z' is a particular label for sections that strongly differ from the rest of the piece or shouldn't be considered part of it (e.g., applause or spoken dialogue).

A prime symbol (') indicates a section similar to a previous one but has significant changes (e.g., transposition or mode change).

Given the hierarchical nature of music, annotators should mark musical similarity on at least two scales. The large-scale sections (A, B, C, ...) should be divided into subsections using lowercase letters (a, b, c, ...). All boundaries marked by an uppercase letter should also be marked by a lowercase letter. Lowercase letters have significance across large-scale sections, indicating the same musical idea even if used in both sections "A" and "B".

Musical Function: It's crucial to correctly identify different sections of the song and label them accordingly. These labels include verse, chorus (or refrain), bridge, intro, outro, and solo. Each of these has a specific role and characteristic:

Verse: This is where the lyrics often change but the melody remains the same.
Chorus (Refrain): This part contrasts with the verse and is repeated more strictly.
Bridge: This section contrasts with the verse and chorus and often serves as a transition.
Intro: This part leads into the rest of the piece.
Outro: This part signals the end of the piece.
Solo: This is when a single instrument or voice is highlighted.
If the song doesn't strictly follow traditional pop song structure (verse, chorus, etc.), you might use labels like main theme and secondary theme.

Leading Instrumentation: In polyphonic music, you must identify and label the instrument or voice that carries the main melody in each segment. If there's no clear lead, that's fine; if two instruments seem to share the lead, you can label both. For instance, in a rock song, you might label segments with the lead vocal as "vocal" and segments where the guitar carries the melody as "guitar" or "electric guitar".

Remember, your analysis should aim to represent the structure and prominent features of the song accurately. The more detail you can provide, the better. Be consistent in labeling and descriptions, and use a standard music analysis vocabulary to ensure others understand your work.

### 2023-05-12 introduction/Subsections/Motivation.tex
Updates from Overleaf

@@

I have consistently been inspired by musicians and researchers who strive to uncover musical ground truth within an existing tonal paradigm, such as Heinrich Schenker and his school of thought, or by challenging it, exemplified by Arnold Schoenberg or George Russell \cite{LydianRussell}. These approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth, development, and further understanding of tonal paradigms.

>>>>

On top of that, I argue that MIR research needs to incorporate a more balanced approach that considers the interdisciplinary nature of music and the importance of other domains beyond DSP.

@@

I have consistently been inspired by musicians and researchers who strive to uncover musical ground truth within an existing tonal paradigm, such as Heinrich Schenker and his school of thought \cite{Komar1959SchenkersStructure}, particularly by challenging it, exemplified by Arnold Schoenberg \cite{Samson1974SchoenbergsMusic}, George Russell \cite{LydianRussell} or Ernst Levy \cite{LevyAHarmony}. These approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth, development, and further understanding tonal and atonal paradigms.

### 2023-05-12 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

@@

In this section, we discuss the concepts of known and unknown invariance in pattern recognition. Invariance refers to the property of a model or algorithm being robust to certain transformations or changes in the input data. In other words, an invariant model should produce the same or similar output for different instances of the same object, even if they have undergone some changes.

Known invariance: Known invariances are transformations or changes that the model or algorithm is explicitly designed to be robust against. This can include the input data's rotations, translations, scaling, or other known perturbations. By incorporating these invariances into the model's design, the algorithm is better equipped to generalize and recognize patterns in the presence of these transformations.

Unknown invariance: Unknown invariances are transformations or changes the model has not been explicitly designed to handle. However, it still manages to be robust due to its learning capabilities. These invariances can emerge from the data and may not be known a priori. For example, a deep learning model might automatically learn to be invariant to certain lighting conditions or other factors not explicitly specified in its design.

@@

The SampleCNN model \cite{CLMR2021} takes in 1D input data with a single channel. The first layer is a 1D convolutional layer with a kernel size of 3, a stride of 3, and 128 output channels. It is followed by batch normalization and ReLU activation.

@@

%\input{figures/neural networks/pipeline}

>>>>

\input{figures/neural networks/pipeline}

@@

In pattern recognition, \textit{invariance} refers to the property of a model to be robust to certain transformations or changes in the input data. In other words, an invariant model should produce the same or similar output for different instances of the same object, even if they have undergone some changes.

\begin{definition}[Known invariance]
Known invariances are transformations or changes that the model or algorithm is explicitly designed to be robust against. This can include the input data's rotations, translations, scaling, or other known perturbations. By incorporating these invariances into the model's design, the algorithm is better equipped to generalize and recognize patterns in the presence of these transformations.
\end{definition}

\begin{definition}[Unknown invariance]

Unknown invariance: Unknown invariances are transformations or changes the model has not been explicitly designed to handle. However, it still manages to be robust due to its learning capabilities. These invariances can emerge from the data and may not be known a priori. For example, a deep learning model might automatically learn to be invariant to certain lighting conditions or other factors not explicitly specified in its design.
\end{definition}

@@

The SampleCNN model \cite{Lee2018SampleCNN:Classification} takes in 1D input data with a single channel. The first layer is a 1D convolutional layer with a kernel size of 3, a stride of 3, and 128 output channels. It is followed by batch normalization and ReLU activation.

@@
